---
title: "Births from 2000 to 2014"
output: github_document
---

Data for this example is taken from [this repository](https://github.com/fivethirtyeight/data/blob/master/births/US_births_2000-2014_SSA.csv). This example uses sqlite 3 but explains what to change to use other methods like Azure SQL Database (used by SQL Server Management Studio). Really only one line in the following R code block needs to be changed to use other methods.

# Connecting to the SQL Server

```{r, .numberLines, .lineAnchors}
library('DBI')
con <- dbConnect(RSQLite::SQLite(), dbname = "rmarkdown-and-sql-built-births-display.db")
birth_data <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv')
dbWriteTable(con, "births2000_2014", birth_data, overwrite = TRUE)
```

# Displaying Table in SQL Code Block

```{sql, connection = con}
SELECT year, month, date_of_month, births FROM births2000_2014
WHERE births > 9000;
```

#Disconnecting from the SQL Server

```{r}
dbDisconnect(con)
```
