---
title: "Births from 2000 to 2014"
output: github_document
---

Data for this example is taken from [this repository](https://github.com/fivethirtyeight/data/blob/master/births/US_births_2000-2014_SSA.csv). This example uses sqlite 3 but explains what to change to use other methods like Azure SQL Database (used by SQL Server Management Studio). Really only one line in the following R code chunk needs to be changed to use other methods. This github flavored markdown file is generated from [this rmarkdown file](rmarkdown-and-sql-built-births-display.rmd).

# Connecting to the SQL Server

```{r, .numberLines, .lineAnchors}
library('DBI')
library('reshape')
con <- dbConnect(RSQLite::SQLite(), dbname = "rmarkdown-and-sql-built-births-display.db")
birth_data <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv')
melted_birth_data <- melt(birth_data, id = "year", measure = "births")
birth_data_by_year <- as.data.frame(cast(melted_birth_data, year~variable, sum))
dbWriteTable(con, "births2000_2014", birth_data_by_year, overwrite = TRUE)
```

For this example to work one must have DBI, reshape and RSQLite R packages installed. One also needs to have RSQLite for one's OS installed too. If one is using a different SQL database type then one needs to have that installed on their OS and have the appropriately DBI supported R package for that type installed as well ([see this for supported DBI backend packages](https://dbi.r-dbi.org/)). One would also have to have the dbConnect call of line 3 of the above R code chunk changed to the call appropriate for the given database type.

## Example with Azure SQL Database

Change line 3 of the above R code chunk to the following:

```{r sql-connect, eval = FALSE}
con <- dbConnect(AzureKusto::AzureKusto(), server="https://azure-database-url.net",
database="desired-database", tenantid="appropriate-azure-tenantid")
```

. Of course AzureKusto R package must be installed as well.

# Displaying Table with SQL Code Block

```{sql, connection = con}
SELECT * FROM births2000_2014
WHERE births > 4200000;
```

This is merely to show that it is possible to display tables generated from SQL code chunks.
For more customisation one should use R packages within R code chunks like shown in the next section.

# Displaying Data from R Code Chunks

```{r sql-display}
library('ggplot2')
my_birth_data_peak_week_day = list()
my_birth_data_peak_week_day_births = list()
my_birth_data_peak_month = list()
my_birth_data_peak_month_births = list()
my_birth_data_average_births_per_month = list()
my_birth_data_total_births = list()
my_birth_data_year = list()
total_births_by_day_of_week = c(0, 0, 0, 0, 0, 0, 0)
total_births_by_month = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
for (i in 1:length(birth_data$year)) {
	year = birth_data$year[i]
	if (year %in% my_birth_data_year) {
		total_births_by_day_of_week = total_births_by_day_of_week
		total_births_by_month = total_births_by_month
		births = birth_data$births[i]
		week_day = birth_data$day_of_week[i]
		month = birth_data$month[i]
		total_births_by_day_of_week[week_day] = total_births_by_day_of_week[week_day] + births
		total_births_by_month[month] = total_births_by_month[month] + births
		total_births_by_day_of_week <<- total_births_by_day_of_week
		total_births_by_month <<- total_births_by_month
	} else {
		if (i > 1) {
			my_birth_data_peak_week_day = my_birth_data_peak_week_day
			my_birth_data_peak_week_day_births = my_birth_data_peak_week_day_births
			my_birth_data_peak_month = my_birth_data_peak_month
			my_birth_data_peak_month_births = my_birth_data_peak_month_births
			my_birth_data_average_births_per_month = my_birth_data_average_births_per_month
			my_birth_data_total_births = my_birth_data_total_births
			peak_week_day_births = max(total_births_by_day_of_week)
			peak_week_day = which(total_births_by_day_of_week == peak_week_day_births)[1]
			peak_month_births = max(total_births_by_month)
			peak_month = which(total_births_by_month == peak_month_births)[1]
			my_birth_data_peak_week_day <<- append(my_birth_data_peak_week_day, peak_week_day)
			my_birth_data_peak_week_day_births <<- append(my_birth_data_peak_week_day_births, peak_week_day_births)
			my_birth_data_peak_month <<- append(my_birth_data_peak_month, peak_month)
			my_birth_data_peak_month_births <<- append(my_birth_data_peak_month_births, peak_month_births)
			total_births = sum(total_births_by_day_of_week)
			my_birth_data_total_births <<- append(my_birth_data_total_births, total_births)
			my_birth_data_average_births_per_month <<- append(my_birth_data_average_births_per_month, total_births/12)
			total_births_by_day_of_week <<- c(0, 0, 0, 0, 0, 0, 0)
			total_births_by_month <<- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
		}
		my_birth_data_year = my_birth_data_year
		my_birth_data_year <<- append(my_birth_data_year, year)
	}
}
peak_week_day_births = max(total_births_by_day_of_week)
peak_week_day = which(total_births_by_day_of_week == peak_week_day_births)[1]
peak_month_births = max(total_births_by_month)
peak_month = which(total_births_by_month == peak_month_births)[1]
my_birth_data_peak_week_day <<- append(my_birth_data_peak_week_day, peak_week_day)
my_birth_data_peak_week_day_births <<- append(my_birth_data_peak_week_day_births, peak_week_day_births)
my_birth_data_peak_month <<- append(my_birth_data_peak_month, peak_month)
my_birth_data_peak_month_births <<- append(my_birth_data_peak_month_births, peak_month_births)
total_births = sum(total_births_by_day_of_week)
my_birth_data_total_births <<- append(my_birth_data_total_births, total_births)
my_birth_data_average_births_per_month <<- append(my_birth_data_average_births_per_month, total_births/12)
my_birth_data = data.frame(year = unlist(my_birth_data_year), peak_week_day = unlist(my_birth_data_peak_week_day), peak_week_day_births = unlist(my_birth_data_peak_week_day_births), peak_month = unlist(my_birth_data_peak_month), peak_month_births = unlist(my_birth_data_peak_month_births), average_births_per_month = unlist(my_birth_data_average_births_per_month), total_births = unlist(my_birth_data_total_births))
knitr::kable(my_birth_data)
ggplot(my_birth_data, aes(x = year, y = total_births)) + geom_line() + geom_point() + labs(x = "YEAR", y = "TOTAL BIRHTS") + scale_x_continuous(breaks = my_birth_data$year, labels = as.character(my_birth_data$year))
```

# Disconnecting from the SQL Server

```{r}
dbDisconnect(con)
```
